---
title: "simpleRegress"
author: "Cesar Guillen"
date: "Summer 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```

## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

```{r}
simp <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simp)
summary.lm(simp)
```

When we create another boxplot to compare which variable would be better to compare a length of a bear, we see that by doing the comparison with the NECK we get a lower Adjusted R-squared value than doing a comparison with AGE. Yes, it is a small increase, but we can also see the difference in the appearance of the boxplot. 

```{r}
simp <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```

The Adjusted R- squared value for using HEADLEN instead of AGE is much larger, meaning that if we would use the variable HEADLEN would give us better predictions for the total length of the bear. We can also see that based on the actual boxplot that the HEADLEN seems to have more unified points that are more uniformed to the line. 

