---
title: "simpleRegress"
author: "Cesar Guillen"
date: "Summer 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```

## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

```{r}
simp <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simp)
summary.lm(simp)
```

When we create another boxplot to compare which variable would be better to compare a length of a bear, we see that by doing the comparison with the NECK we get a lower Adjusted R-squared value than doing a comparison with AGE. Yes, it is a small increase, but we can also see the difference in the appearance of the boxplot. 

```{r}
simp <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```

The Adjusted R- squared value for using HEADLEN instead of AGE is much larger, meaning that if we would use the variable HEADLEN would give us better predictions for the total length of the bear. We can also see that based on the actual boxplot that the HEADLEN seems to have more unified points that are more uniformed to the line. 

```{r}
simp <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```

If we would use HEADWTH then we would get a lower Adjusted R-squared meaning it would not be as useful as using some of the other variables we are given. Even the residual standard error is much larger than the other two variables we have used, which can mean we would get more errors than the other predictions. 

```{r}
simp <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```

While using the variable CHEST we see that it is better than using the variable AGE, where it has a greater adjusted r- squared value, but it is not as great if we compare it to the value of the adjusted r- squared, where CHEST has a smaller value than HEADLEN. So in comparison HEADLEN is still the better variable to use to find the better predictions of total length. Both also have a medium sized residual standard error, but HEADLEN is still the one with a smaller variable. 



